Temperature and moisture effects on greenhouse gas emissions from deep active-layer boreal soils

### 2016-03-01 manuscript/cpcrw_manuscript.Rmd
Added tentative UGA text, and flux summary results

# Run the analysis script, which will leave a lot of figures, tables, and data in memory for us
setwd("../")
source("R/5-analyze.R")
```

@@

The magnitude, timing, and form-in particular as methane (CH~4~) or carbon dioxide (CO~2~)-of such a feedback remain highly uncertain {Schuur, 2015 #4319}. While northern soils hold enormous quantities {Tarnocai, 2009 #2231} of potentially mineralizable soil organic carbon (SOC), vegetation succession dynamics promote permafrost resilience to even large temperature changes {Jorgenson, 2010 #2627}. Such dynamics may however be disrupted by increased fire disturbance {Johnstone, 2010 #2465}. In addition, the stability of SOC is itself highly uncertain, as it depends on soil temperature and moisture, the ages of and ratio between the carbon and nitrogen pools {Weiss, 2015 #4317}, and its protection from competent microorganisms, enzymes, and resources {Bailey, 2012 #3624}, whether by organomineral sorption, chemical lability, or physical location {Schmidt, 2011 #3135}.

@@

The incubation experiment concluded on 9 December 2015, following the final CO~2~ and CH~4~ readings. Each soil core was maintained at the treatment-dependent temperature and moisture content (by mass) until removed for destructive sampling, December 14-18, 2015. Sub-samples were collected and composited throughout each soil core for dissolved organic carbon analysis (110 ± 24 g dry mass equivalent) and dry-mass calculations (~28 g each). The remaining core material was air-dried and separated into particles (>2 mm diameter) and soil (≤2 mm) using a U.S. Standard Test Sieve No. 10 (Fisherbrand,  Pittsburg, PA, USA). The dry mass and volume of soil were used in calculations of gravimetric and volumetric soil moisture content, respectively {Gardner, 1986 #4345}. Soil volume was caclulated as the total core volume minus the volume of particles >2 mm diameter, with the latter determined by water displacement. Air-dried soil and sub-samples stored at -20 °C were sent to the Agricultural and Environmental Services Laboratory at the University of Georgia Extension in February 2016 for total C, N, and dissolved organic C. TODO: get UGA analytical methods.

@@

* Soil C, N, DOC means

* Gas fluxes over incubation (Figures 2 and 3)

@@

* Cumulative fluxes (Figure 4)

@@

* CO2:CH4 ratios

>>>>

# Run the main analysis script, which will leave figures, tables, and data 
# in memory for us. Need to temporarily change the working directory to do so.
olddir <- getwd()
setwd("../")
source("R/5-analyze.R")
setwd(olddir)
```

@@

The magnitude, timing, and form-in particular as methane (CH~4~) or carbon dioxide (CO~2~)-of such a feedback remain highly uncertain {Schuur, 2015 #4319}. While northern soils hold enormous quantities {Tarnocai, 2009 #2231} of potentially mineralizable soil organic carbon (SOC), vegetation succession dynamics promote permafrost resilience to even large temperature changes {Jorgenson, 2010 #2627}. Such dynamics may however be disrupted by increased fire disturbance {Johnstone, 2010 #2465}. In addition, the stability of SOC is itself highly uncertain, as it depends on soil temperature and moisture, the ages of and ratio between the carbon (C) and nitrogen (N) pools {Weiss, 2015 #4317}, and its protection from competent microorganisms, enzymes, and resources {Bailey, 2012 #3624}, whether by organomineral sorption, chemical lability, or physical location {Schmidt, 2011 #3135}.

@@

The incubation experiment concluded on 9 December 2015, following the final CO~2~ and CH~4~ readings. Each soil core was maintained at the treatment-dependent temperature and moisture content (by mass) until removed for destructive sampling, December 14-18, 2015. Sub-samples were collected and composited throughout each soil core for dissolved organic carbon analysis (110 ± 24 g dry mass equivalent) and dry-mass calculations (~28 g each). The remaining core material was air-dried and separated into particles (>2 mm diameter) and soil (≤2 mm) using a U.S. Standard Test Sieve No. 10 (Fisherbrand,  Pittsburg, PA, USA). The dry mass and volume of soil were used in calculations of gravimetric and volumetric soil moisture content, respectively {Gardner, 1986 #4345}. Soil volume was caclulated as the total core volume minus the volume of particles >2 mm diameter, with the latter determined by water displacement. 
Air-dried soil and sub-samples stored at -20 °C were sent to the Agricultural and Environmental Services Laboratory at the University of Georgia Extension in February 2016 for total C, N, and dissolved organic C. Samples were combusted in an oxygen atmosphere at 1350 °C, converting elemental C and N into gaseous form. Infrared cells were then used to to determine C content, and a thermal conductivity cell to determine N. TODO: confirm UGA analytical methods and exact instruments.

@@

After removal of known problematic data, and exclusion of statistical outliers, CO~2~ flux rates measured during the incubation ranged from `r pn(fluxsummary["flux_µgC_gC_day_min", "CO2"]/1000, 3)` mg C g C^-1^ day^-1^ (`r pn(fluxsummary["flux_µgC_g_day_min", "CO2"]/1000, 3)` mg C g soil^-1^ day^-1^) to a maximum of `r pn(fluxsummary["flux_µgC_gC_day_max", "CO2"]/1000, 3)` (`r pn(fluxsummary["flux_µgC_g_day_max", "CO2"]/1000, 3)`), with a mean of `r pn(fluxsummary["flux_µgC_gC_day_mean", "CO2"]/1000, 3)` (`r pn(fluxsummary["flux_µgC_g_day_mean", "CO2"]/1000, 3)`). CH~4~ rates ranged from `r pn(fluxsummary["flux_µgC_gC_day_min", "CH4"]*1000, 3)` ng C g C^-1^ day^-1^ (`r pn(fluxsummary["flux_µgC_g_day_min", "CH4"]*1000, 3)` ng C g soil^-1^ day^-1^) to a maximum of `r pn(fluxsummary["flux_µgC_gC_day_max", "CH4"]*1000, 3)` (`r pn(fluxsummary["flux_µgC_g_day_max", "CH4"]*1000, 3)`), with a mean of `r pn(fluxsummary["flux_µgC_gC_day_mean", "CH4"]*1000, 3)` (`r pn(fluxsummary["flux_µgC_g_day_mean", "CH4"]*1000, 3)`). Gas fluxes varied considerably over the course of the incubation (**Figures 2 and 3**).

@@

@@
